Analysis paper

### 2021-06-24 analysis/analysis.Rmd
update plots, simplify code

# define colors for deaths and cases
colcases <- brewer.pal(name = "Accent", n = 5)[-c(1, 4)]
coldeaths <- brewer.pal(name = "Accent", n = 5)[-4]

@@

## Visualsiation of daily data

```{r}
dailytruth_data %>%
  dplyr::filter(target_end_date >= min(forecast_dates$full_hub_period), 
                target_end_date <= max(forecast_dates$full_hub_period)) %>%
  ggplot(aes(y = true_value, x = target_end_date)) + 
  geom_point(size = 0.3, alpha = 0.4) + 
  geom_line(size = 0.5, alpha = 0.4) + 
  geom_line(data = truth_data, aes(y = true_value / 7)) + 
  facet_grid(target_type ~ location_name, scales = "free") + 
  labs(y = "Observed values", x = "Date") + 
  scale_x_date(limits = c(min(forecast_dates$full_hub_period) - 28, 
                            max(forecast_dates$full_hub_period)))
  theme_minimal()

@@

## Visualisation for 1, 2, 3, 4 week ahead forecasts and scores

### Plots that visualise truth vs. predictions

```{r truth-vs-forecasts, fig.height=12, fig.width=15}
# create helper function for the plot ------------------------------------------
truth_and_pred_plot <- function(df, ttype = "case") {
  p <- df %>%
    dplyr::filter(target_type == ttype) %>%
    ggplot(aes(y = true_value, x = target_end_date)) + 
    geom_linerange(aes(ymin = `0.25`, ymax = `0.75`, color = model), 
                   size = 1.7,
                   alpha = 1,
                   position = position_dodge(width = 4)) + 
    # geom_linerange(aes(ymin = `0.025`, ymax = `0.975`, color = model), 
    #                size = 1.7,
    #                alpha = 0.1,
    #                position = position_dodge(width = 4)) +   
    geom_point(data = unfiltered_data[target_type == ttype], 
               aes(shape = classification)) + 
    geom_line(data = unfiltered_data[target_type == ttype]) + 
    theme_minimal() + 
    facet_wrap(~ location_target, scales = "free", ncol = 2) + 
    theme(legend.position = "bottom") + 
    # theme(panel.background = element_rect(fill = "#e1f0fe")) + 
    scale_x_date(limits = c(min(forecast_dates$full_hub_period) + 5, 
                            max(forecast_dates$full_hub_period)))
  return(p)
}

# prepare data for plotting ----------------------------------------------------
# - load ufiltered data and pivot wider
# - add trend classification by target_end_date 
#   (and remove classification by forecast_date that is already present)

df <- unfiltered_data[model %in% regular_models &
                        quantile %in% c(0.025, 0.25, 0.5, 0.75, 0.975)] %>%
  dcast(... ~ quantile, value.var = "prediction")
df <- merge(
  df[, c("classification", "speed") := NULL], 
  copy(epitrend)[, forecast_date := NULL], 
  by = c("target_end_date", "location_name", "target_type")
)

# create plots for deaths and cases --------------------------------------------

for (i in 1:4) {
  # plot for cases
  truth_vs_forecast_cases <- df[grepl(i, target)] %>%
    truth_and_pred_plot(ttype = "case") + 
    scale_color_manual(values = colcases) + 
    labs(y = "Cases", x = "")

  filename <- here("analysis", "plots", 
                   paste0("truth-and-", i, "-wk-ahead_cases"))
  ggsave(paste0(filename, ".png"), plot = truth_vs_forecast_cases)

  saveRDS(truth_vs_forecast_cases, 
          paste0(filename, ".RDS"))

  # plot for deaths
  truth_vs_forecast_deaths <- df[grepl(i, target)] %>%
  truth_and_pred_plot(ttype = "death")  + 
  scale_color_manual(values = coldeaths) + 
  annotate('rect', xmin = min(forecast_dates$full_hub_period) + 14,
                xmax = max(forecast_dates$death_not_scored) + 14,
                ymin = 0,
                ymax = Inf, fill = "grey10", alpha = 0.1) + 
    labs(y = "Deaths", x = "")

  filename <- here("analysis", "plots", 
                   paste0("truth-and-", i, "-wk-ahead_deaths"))
  ggsave(paste0(filename, ".png"), plot = truth_vs_forecast_deaths)

  saveRDS(truth_vs_forecast_deaths, 
          paste0(filename, ".RDS"))
}
```

### Plots that visualise the achieved scores over time

```{r scores-over-time, fig.height=10, fig.width=15}
# score forecasts --------------------------------------------------------------
# obtain scores and then do pairwise comparisons between models
summarise_by <- c("model", "horizon", "location_name", "target_type", 
                   "forecast_date", "target_end_date", "location_target", 
                   "classification")
scores_target_forecastdate <- eval_forecasts(
  data = unfiltered_data[model %in% regular_models],
  summarise_by = summarise_by
)

# pairwise <- pairwise_comparison(scores_target_forecastdate, 
#                                 summarise_by = summarise_by, 
#                                 baseline = "Baseline")
# pairwise <- pairwise[compare_against == "Baseline"]

# create helper function to plot forecasts -------------------------------------
scores_over_time_plot <- function(scores, y = "mean_scores_ratio") {
  p <- scores %>%
  ggplot(aes(x = as.Date(target_end_date), y = get(y))) + 
  geom_line(aes(colour = model)) + 
  geom_point(aes(shape = classification, colour = model)) + 
  theme_minimal() + 
  facet_wrap( ~ location_target, ncol = 2, scales = "free_y") +
  theme(legend.position = "bottom") +
  # geom_rect(aes(xmin = min(forecast_dates$christmas) + 14,
  #               xmax = max(forecast_dates$christmas) + 14,
  #               ymin = 0,
  #               ymax = Inf), fill = "lightgrey", alpha = 0.002) +
  # theme(panel.background = element_rect(fill = "#e1f0fe")) +
  # scale_y_log10() + 
  scale_x_date(limits = c(min(forecast_dates$full_hub_period) + 5, 
                          max(forecast_dates$full_hub_period)))
  return(p)
}

# create plots for all horizons ------------------------------------------------
for (i in 1:4) {
  scores_over_time_case <- scores_target_forecastdate[horizon == i & target_type == "case"] %>%
    scores_over_time_plot(y = "interval_score") + 
    scale_color_manual(values = colcases) + 
    labs(y = "WIS rel. to baseline", x = "")

  ggsave(here("analysis", "plots", paste0("score-over-time-cases-", i, ".png")), 
         plot = scores_over_time_case)

  saveRDS(scores_over_time_case, 
          here("analysis", "plots", paste0("score-over-time-cases-", i, ".RDS")))

  scores_over_time_deaths <- scores_target_forecastdate[horizon == i & target_type == "death"] %>%
    scores_over_time_plot(y = "interval_score") + 
    scale_color_manual(values = coldeaths) + 
    annotate('rect', xmin = min(forecast_dates$full_hub_period) + 14,
             xmax = max(forecast_dates$death_not_scored) + 14,
             ymin = 0,
             ymax = Inf, fill = "grey10", alpha = 0.1) + 
    labs(y = "WIS rel. to baseline", x = "Date")

  ggsave(here("analysis", "plots", paste0("score-over-time-deaths-", i, ".png")), 
         plot = scores_over_time_deaths)

  saveRDS(scores_over_time_deaths, 
          here("analysis", "plots", paste0("score-over-time-deaths-", i, ".RDS")))

}
```

### Combine the two plots

```{r fig.height=15, fig.width=10}
horizons <- 1:4

for (horizon in horizons) {
  truth_vs_pred_cases <- readRDS(file = here("analysis", "plots", 
                                             paste0("truth-and-", horizon, 
                                                    "-wk-ahead_cases.RDS")))

  truth_vs_pred_deaths <- readRDS(file = here("analysis", "plots", 
                                              paste0("truth-and-", horizon, 
                                                     "-wk-ahead_deaths.RDS")))

  scores_over_time_cases <- readRDS(file = here("analysis", "plots", 
                                                paste0("score-over-time-cases-", 
                                                       horizon,".RDS")))

  scores_over_time_deaths <- readRDS(file = here("analysis", "plots", 
                                                 paste0("score-over-time-deaths-", 
                                                        horizon,".RDS")))

  truth_vs_pred_cases / scores_over_time_cases / truth_vs_pred_deaths / scores_over_time_deaths + 
    plot_layout(guides = "collect") & 
    plot_annotation(tag_levels = 'A') &
    theme(legend.position = 'bottom')

  ggsave(here("analysis", "plots", paste0("figure-forecasts-", horizon, ".png")), width = 10, height = 12)
}

@@

## Summary table with scores

```{r}
# create helper function to make table
make_score_table <- function(scores, 
                             group = TRUE, 
                             full_width = TRUE) {
  # copy data, change names, round column values and reduce to sign. digits
  scores <- as.data.table(scores)
  setnames(scores, 
           old = c("horizon", "model", "target_type", "interval_score", "aem", 
                   "relative_skill", "scaled_rel_skill", "sharpness", 
                   "underprediction", "overprediction", "bias", 
                   "50", "90", "location_target", "target_phase"), 
           new = c("Horizon", "Model", "Target", "WIS", "Abs. error", 
                   "rel. skill", "WIS - rel.", "Sharpness", 
                   "Underpred.", "Overpred.", "Bias", 
                   "50%-Cov.", "90%-Cov.", "Target", "Target"), 
           skip_absent = TRUE)

@@

### Create tables

```{r }
# create scores ----------------------------------------------------------------
# calculate coverage independently and merge with other scores
coverage <- eval_forecasts(
  filtered_data[model %in% regular_models], 
  summarise_by = c("horizon", "model", "target_type", "range"), 
)[range %in% c(50, 90), 
                   .(model, target_type, coverage, range, horizon)] %>%
  dcast(formula = ... ~ range, value.var = "coverage")

df <- eval_forecasts(
  filtered_data[model %in% regular_models], 
  summarise_by = c("model", "horizon", "target_type"), 
  compute_relative_skill = FALSE,
  sd = TRUE, 
  quantiles = 0.5) %>%
  merge(coverage, all.x = TRUE, by = c("model", "horizon", "target_type"))

@@

# tables for 1-2 weeks ahead and 3-4 weeks ahdead
table <- make_score_table(df[horizon <= 2]) 
save_kable(table, file = here("analysis", "plots", "table_scores_2_ahead.png"))

table <- make_score_table(df[horizon > 2]) 
save_kable(table, file = here("analysis", "plots", "table_scores_4_ahead.png"))

@@

### Create table for all median ensembles

```{r }
# create scores ----------------------------------------------------------------
# calculate coverage independently and merge with other scores
coverage <- eval_forecasts(
  filtered_data[model %in% ensemble_models[!grepl("mean", ensemble_models)]], 
  summarise_by = c("horizon", "model", "target_type", "range"), 
)[range %in% c(50, 90), 
                   .(model, target_type, coverage, range, horizon)] %>%
  dcast(formula = ... ~ range, value.var = "coverage")

df <- eval_forecasts(
  filtered_data[model %in% ensemble_models[!grepl("mean", ensemble_models)]], 
  summarise_by = c("model", "horizon", "target_type"), 
  compute_relative_skill = FALSE, 
  sd = TRUE, 
  quantiles = 0.5,
) %>%
  merge(coverage, all.x = TRUE, by = c("model", "horizon", "target_type")) 
del_cols <- names(df)[(grepl("_sd", names(df)) | grepl("0.5", names(df))) &
                        !(grepl("interval", names(df)))] %>%
  c("coverage_deviation")
df[, (del_cols) := NULL]

# tables for 1-2 weeks ahead and 3-4 weeks ahdead
table <- make_score_table(df[horizon <= 2]) 
save_kable(table, file = here("analysis", "plots", "table_median-ensemble_scores_2_ahead.png"))

table <- make_score_table(df[horizon > 2]) 
save_kable(table, file = here("analysis", "plots", "table_median-ensemble_scores_4_ahead.png"))

### Create table for all mean ensembles

```{r }
# create scores ----------------------------------------------------------------
# calculate coverage independently and merge with other scores
coverage <- eval_forecasts(
  filtered_data[model %in% ensemble_models[grepl("mean", ensemble_models)]], 
  summarise_by = c("horizon", "model", "target_type", "range"), 
)[range %in% c(50, 90), 
                   .(model, target_type, coverage, range, horizon)] %>%
  dcast(formula = ... ~ range, value.var = "coverage")

df <- eval_forecasts(
  filtered_data[model %in% ensemble_models[grepl("mean", ensemble_models)]], 
  summarise_by = c("model", "horizon", "target_type"), 
  compute_relative_skill = FALSE, 
  sd = TRUE, 
  quantiles = 0.5,
) %>%
  merge(coverage, all.x = TRUE, by = c("model", "horizon", "target_type")) 
del_cols <- names(df)[(grepl("_sd", names(df)) | grepl("0.5", names(df))) &
                        !(grepl("interval", names(df)))]
del_cols <- c(del_cols, "coverage_deviation")
df[, (del_cols) := NULL]

# tables for 1-2 weeks ahead and 3-4 weeks ahdead
table <- make_score_table(df[horizon <= 2]) 
save_kable(table, file = here("analysis", "plots", "table_mean-ensemble_scores_2_ahead.png"))

table <- make_score_table(df[horizon > 2]) 
save_kable(table, file = here("analysis", "plots", "table_mean-ensemble_scores_4_ahead.png"))

```{r}
# calculate median and sd of the WIS -------------------------------------------

df <- eval_forecasts(
  filtered_data[model %in% regular_models], 
  metric = "interval_score",
  summarise_by = c("model", "horizon", "target_type"), 
  sd = TRUE, 
  quantile = 0.5
)

df <- df[, .(model, horizon, target_type, 
             `WIS - mean` = interval_score, 
             `WIS - median` = interval_score_0.5, 
             `WIS - sd` = interval_score_sd)] %>%
    dcast(... ~ ., value.var = c("WIS - mean", "WIS - median", "WIS - sd"))

setcolorder(df, c("target_type", "horizon", "model"))

table <- df[order(target_type, horizon, model)] %>%
  make_score_table()

save_kable(table, 
           file = here("analysis", "plots", "table-wis-sd-median.png"))

@@

# for horizon <= 2
labels <- df[horizon <= 2][order(target_phase, horizon)]$target_phase
table <- make_score_table(df[horizon <= 2, !"target_type"], group = FALSE) %>%
  pack_rows(index = table(labels), indent = FALSE)
save_kable(table, 
           file = here("analysis", "plots", "table-phases-2.png"))

# for horizon > 2
labels <- df[horizon > 2][order(target_phase, horizon)]$target_phase
table <- make_score_table(df[horizon > 2, !"target_type"], group = FALSE) %>%
  pack_rows(index = table(labels), indent = FALSE)
save_kable(table, 
           file = here("analysis", "plots", "table-phases-4.png"))

# 
# horizons = 2
# for (tar in targets) {
#   for (i in horizons) {
#     name <- paste0(tar, "-", i)
#     
#     tmp_df <- df[target_phase == tar & horizon == i, 
#                  .(model, interval_score, sharpness,
#                    target_phase, horizon, 
#                    underprediction, overprediction, 
#                    `WIS - median` = interval_score_0.5, 
#                    `WIS - sd` = interval_score_sd)]
#     
#     tables[[name]] <- make_score_table(tmp_df, 
#                                        group = FALSE, 
#                                        full_width = FALSE) %>%
#       pack_rows(tar, 1, nrow(tmp_df), indent = FALSE)
#     
#     save_kable(tables[[name]],
#                file = here("analysis", "plots",
#                            paste0("table-", gsub(" ", "", name), ".png")))
#   }
# }

@@

```{r fig.height=10, fig.width=15}
# 
# coverage_tables <- eval_forecasts(
#   filtered_data[grepl("2", target)], 
#   summarise_by = c("model", "target_phase", "classification", "range", "horizon"), 
#   compute_relative_skill = TRUE, 
#   baseline = "Baseline"
# )[range %in% c(50, 90), 
#                    .(model, target_phase, coverage, classification, range)] %>%
#   dcast(formula = ... ~ range, value.var = "coverage")
# 
# scores_tables <- eval_forecasts(
#   filtered_data[grepl("2", target)], 
#   summarise_by = c("model", "target_phase", "classification", "horizon"), 
#   compute_relative_skill = TRUE, 
#   baseline = "Baseline"
# )
# 
# tbl <- scores_tables %>%
#   merge(coverage_tables, all.x = TRUE, by = c("model", "target_phase", "classification")) 
# 
#  tbl <- tbl[, lapply(.SD, FUN = function(x) {
#     if (is.numeric(x)) {
#       return(round(signif(x, 3), 2))
#     } else {
#       return(x)
#     }
#   })]
# 
# tables <- list()
# table_dfs <- list()
# targets <- unique(tbl$target_phase)
# 
# for (tar in targets) {
#   tables[[tar]] <- make_score_table(tbl[target_phase == tar])
# }
# 
# for (tar in targets) {
#   table_dfs[[tar]] <- tbl[target_phase == tar, 
#                           .("Model" = model, 
#                             # "Target" = target_phase,
#                             "WIS" = interval_score,
#                             # "Rel. skill" = scaled_rel_skill, 
#                             "Sharpn." = sharpness, 
#                             "Underp.." = underprediction, 
#                             "Overp." = overprediction)#, 
#                             # "Bias" = bias,
#                             # "50%-Cov." = `50`, 
#                             # "90%-Cov." = `90`)
#                           ]
# }
# 
# 
# # make plots with tables
# library(gridExtra)
# 
# tbl_plt <- grid.arrange(tableGrob(table_dfs[[1]]), 
#              tableGrob(table_dfs[[2]]), 
#              tableGrob(table_dfs[[3]]), 
#              tableGrob(table_dfs[[4]]), 
#              tableGrob(table_dfs[[5]]), 
#              tableGrob(table_dfs[[6]]), 
#              nrow = 2) 
# 
# tbl_plt <- ggplotify::as.ggplot(tbl_plt)
# 
# tbl_plt
# 
# ggsave("plots/distribution_scores_table.png")
# 
# library(gt)
# 
# tbl[, 
#     .("Model" = model, 
#       "Target" = target_phase,
#       "WIS" = interval_score,
#       # "Rel. skill" = scaled_rel_skill, 
#       "Sharpn." = sharpness, 
#       "Underp.." = underprediction, 
#       "Overp." = overprediction)#, 
#     # "Bias" = bias,
#     # "50%-Cov." = `50`, 
#     # "90%-Cov." = `90`)
# ] %>%
#   gt(rowname_col = "Model", 
#      groupname_col = "Target") 
#

@@

scores <- eval_forecasts(filtered_data[model %in% regular_models], 
                         summarise_by = c("model", "target_type", "horizon"), 
                         quantiles = 0.5, 
                         sd = TRUE)

# mean performance
p_mean <- scores %>%
  ggplot(aes(x = horizon, y = interval_score, group = model, color = model)) + 
  geom_point() + 
  geom_line() + 
  facet_wrap(~ target_type, scales = "free_y") + 
  theme_minimal() + 
  theme(legend.position = "bottom") + 
  scale_color_manual(values = coldeaths) + 
  labs(y = "mean WIS", x = NULL)

# ggsave(here("analysis", "plots", "aggregate-performance-mean.png"))

>>>>

# define colors for deaths and cases
colcases <- brewer.pal(name = "Accent", n = 8)[-c(1, 4)]
coldeaths <- brewer.pal(name = "Accent", n = 8)[-4]

@@

--------------------------------------------------------------------------------
## Plot with Visualsiation of daily data

```{r}
dailytruth_data %>%
  dplyr::filter(target_end_date >= min(forecast_dates$full_hub_period), 
                target_end_date <= max(forecast_dates$full_hub_period)) %>%
  ggplot(aes(y = true_value, x = target_end_date)) + 
  geom_point(size = 0.3, alpha = 0.4) + 
  geom_line(size = 0.5, alpha = 0.4) + 
  geom_line(data = truth_data, aes(y = true_value / 7)) + 
  facet_grid(target_type ~ location_name, scales = "free") + 
  labs(y = "Observed values", x = "Date") + 
  scale_x_date(limits = c(min(forecast_dates$full_hub_period) - 28, 
                            max(forecast_dates$full_hub_period))) + 
  theme_minimal()

@@

--------------------------------------------------------------------------------
## Plot with Visualsiation of x week ahead forecasts and scores

```{r truth-vs-forecasts, fig.height=12, fig.width=15}

# create helper functions for the plot 
# ------------------------------------------------------------------------------
# truth_and_pred_plot = helper function for the visualisation truth and forecast
# scores_over_time_plot = helper function for plotting scores over time
# forecasts_and_scores = function that organises creation of the plots

truth_and_pred_plot <- function(df, ttype = "case") {

  trend <- 
    copy(epitrend)[, location_target := paste0(str_to_sentence(target_type), 
                                               "s", " in ", location_name)]
  p <- df %>%
    dplyr::filter(target_type == ttype) %>%
    ggplot(aes(y = true_value, x = target_end_date)) + 
    geom_linerange(aes(ymin = `0.25`, ymax = `0.75`, color = model), 
                   size = 1.7,
                   alpha = 1,
                   position = position_dodge(width = 4.6)) + 
    geom_point(data = trend[target_type == ttype],
               aes(shape = classification)) +
    geom_line(data = unfiltered_data[target_type == ttype & horizon == 2]) + 
    theme_minimal() + 
    facet_wrap(~ location_target, scales = "free", ncol = 2) + 
    theme(legend.position = "bottom") + 
    scale_x_date(limits = c(min(forecast_dates$full_hub_period) + 5, 
                            max(forecast_dates$full_hub_period)))
  return(p)
}

scores_over_time_plot <- function(scores, y = "mean_scores_ratio") {
  p <- scores %>%
  ggplot(aes(x = as.Date(target_end_date), y = get(y))) + 
  geom_line(aes(colour = model)) + 
  geom_point(aes(colour = model, shape = classification)) + 
  theme_minimal() + 
  facet_wrap( ~ location_target, ncol = 2, scales = "free_y") +
  theme(legend.position = "bottom") +
  scale_x_date(limits = c(min(forecast_dates$full_hub_period) + 5, 
                          max(forecast_dates$full_hub_period)))
  return(p)
}

forecasts_and_scores <- function(data, 
                                 output_dir = here("analysis", "plots"), 
                                 plotname = "figure-forecasts", 
                                 ensembles = FALSE) {

  if (ensembles) {
    colcases <- c(colcases[1:2], colcases[4], colcases[3])
    coldeaths <- colcases
  }

  # create data.frame for plotting predictions ---------------------------------
  df <- data[quantile %in% c(0.025, 0.25, 0.5, 0.75, 0.975)] %>%
    dcast(... ~ quantile, value.var = "prediction")
  df <- merge(
    df[, c("classification", "speed") := NULL], 
    copy(epitrend)[, c("forecast_date", "true_value") := NULL], 
    by = c("target_end_date", "location_name", "target_type")
  )

  # create data.frame for plotting scores --------------------------------------
  summarise_by <- c("model", "horizon", "location_name", "target_type", 
                    "forecast_date", "target_end_date", "location_target")

  scores_target_forecastdate <- eval_forecasts(
    data = data,
    summarise_by = summarise_by
  )[, forecast_date := as.Date(forecast_date)]

  # copy epitrend data.frame and only keep necessary columns
  epi_tmp <- copy(epitrend)
  epi_tmp[, forecast_date := (as.Date(forecast_date))]
  epi_tmp <- 
    epi_tmp[, .(location_name, forecast_date, classification, target_type)] %>% 
    unique()

  # create plots for all horizons ----------------------------------------------
  for (i in 1:4) {

    # truth plot for cases 
    truth_vs_forecast_cases <- df[grepl(i, target)] %>%
      truth_and_pred_plot(ttype = "case") + 
      scale_color_manual(values = colcases) + 
      labs(y = "Cases", x = "")

    # truth plot for deaths
    truth_vs_forecast_deaths <- df[grepl(i, target)] %>%
      truth_and_pred_plot(ttype = "death")  + 
      scale_color_manual(values = coldeaths) + 
      annotate('rect', xmin = min(forecast_dates$full_hub_period) + 14,
               xmax = max(forecast_dates$death_not_scored) + 14,
               ymin = 0,
               ymax = Inf, fill = "grey10", alpha = 0.1) + 
      labs(y = "Deaths", x = "")

    # scores plot for cases
    scores_over_time_cases <- 
      scores_target_forecastdate[horizon == i & target_type == "case"] %>%
      merge(epi_tmp) %>%
      scores_over_time_plot(y = "interval_score") + 
      scale_color_manual(values = colcases) + 
      labs(y = "WIS", x = "")

    # scores plot for deaths
    scores_over_time_deaths <- 
      scores_target_forecastdate[horizon == i & target_type == "death"] %>%
      merge(epi_tmp) %>%
      scores_over_time_plot(y = "interval_score") + 
      scale_color_manual(values = coldeaths) + 
      annotate('rect', xmin = min(forecast_dates$full_hub_period) + 14,
               xmax = max(forecast_dates$death_not_scored) + 14,
               ymin = 0,
               ymax = Inf, fill = "grey10", alpha = 0.1) + 
      labs(y = "WIS", x = "Date")

    # putting them all together
    truth_vs_forecast_cases / scores_over_time_cases / 
      truth_vs_forecast_deaths / scores_over_time_deaths + 
      plot_layout(guides = "collect") & 
      plot_annotation(tag_levels = 'A') &
      theme(legend.position = 'bottom')

    ggsave(here(output_dir, 
                paste0(plotname, "-", i, ".png")), width = 10, height = 12)

  }
}

# create plots -----------------------------------------------------------------
# regular forecasts
unfiltered_data[model %in% regular_models] %>%
  forecasts_and_scores()

# ensemble variants
models <- c("Hub-ensemble-all", "Crowd forecast", "Hub-ensemble", "Renewal")
unfiltered_data[model %in% models] %>%
  forecasts_and_scores(plotname = "figure-forecasts-ensembles", 
                       ensembles = TRUE)

@@

--------------------------------------------------------------------------------
### alternative visualisation of forecasts

not yet done

```{r}
scores_over_time_plot_components <- function(scores) {

size = 1.8
width = 4

scores %>%
  ggplot(aes(x = forecast_date, 
             group = model, 
             color = model)) +

  geom_linerange(aes(ymin = -0.5*sharpness, ymax = 0.5 * sharpness, alpha = "Sharpness"), 
                 size = size,
                 alpha = 1,
                 position = position_dodge(width = width)) +

  geom_linerange(aes(ymax = -0.5 * sharpness, ymin = -0.5* sharpness - underprediction),
                 size = size,
                 position = position_dodge(width = width)) + 
  geom_linerange(aes(ymax = -0.5 * sharpness, ymin = -0.5* sharpness - underprediction),
                 size = size,
                 colour = "white",
                 size = size - 0.6,
                 alpha = 0.6,
                 position = position_dodge(width = width)) + 
  geom_linerange(aes(ymin = 0.5 * sharpness, ymax = 0.5* sharpness + overprediction),
                 size = size,
                 position = position_dodge(width = width)) + 
  geom_linerange(aes(ymin = 0.5 * sharpness, ymax = 0.5* sharpness + overprediction),
                 size = size,
                 colour = "white",
                 size = size - 0.6,
                 alpha = 0.6,
                 position = position_dodge(width = width)) + 
  theme_minimal() + 
  scale_x_date(limits = c(min(forecast_dates$full_hub_period) + 5, 
                          max(forecast_dates$full_hub_period)))

   scores_target_forecastdate[horizon == i & target_type == "case"] %>%
    scores_over_time_plot_components() + 
    scale_color_manual(values = colcases) + 
    labs(y = "WIS rel. to baseline", x = "")

--------------------------------------------------------------------------------
## Summary tables

# create helper function to make table
# ------------------------------------------------------------------------------
score_table <- function(scores, 
                             group = TRUE, 
                             full_width = TRUE) {
  # copy data, change names, round column values and reduce to sign. digits
  scores <- as.data.table(scores)
  setnames(scores, 
           old = c("horizon", "model", "target_type", "interval_score", "aem", 
                   "relative_skill", "scaled_rel_skill", "sharpness", 
                   "underprediction", "overprediction", "bias", 
                   "50", "90", "location_target", "target_phase"), 
           new = c("Horizon", "Model", "Target", "WIS", "Abs. error", 
                   "rel. skill", "WIS - rel.", "Sharpness", 
                   "Underpred.", "Overpred.", "Bias", 
                   "50%-Cov.", "90%-Cov.", "Target", "Target"), 
           skip_absent = TRUE)

@@

make_table <- function(data, 
                       plotname = "table_scores",
                       output_dir = here("analysis", "plots")) {

  coverage <- eval_forecasts(
    data = data, 
    summarise_by = c("horizon", "model", "target_type", "range"), 
  )[range %in% c(50, 90), 
    .(model, target_type, coverage, range, horizon)] %>%
    dcast(formula = ... ~ range, value.var = "coverage")

  df <- eval_forecasts(
    data, 
    summarise_by = c("model", "horizon", "target_type"), 
    compute_relative_skill = FALSE,
    sd = TRUE, 
    quantiles = 0.5) %>%
    merge(coverage, all.x = TRUE, by = c("model", "horizon", "target_type"))

@@

  # tables for 1-2 weeks ahead and 3-4 weeks ahdead
  table <- score_table(df[horizon <= 2]) 
  save_kable(table, file = here(output_dir, 
                                paste0(plotname, "_", 2, "_ahead.png")))

  table <- score_table(df[horizon > 2]) 
  save_kable(table, file = here(output_dir, 
                                paste0(plotname, "_", 4, "_ahead.png")))
}

# create tables
# ------------------------------------------------------------------------------

filtered_data[model %in% regular_models] %>%
  make_table()

filtered_data[model %in% ensemble_models[grepl("mean", ensemble_models)]] %>%
  make_table(plotname = "table_mean-ensemble_scores")

filtered_data[model %in% ensemble_models[!grepl("mean", ensemble_models)]] %>%
  make_table(plotname = "table_median-ensemble_scores")

@@

```{r}
# # calculate median and sd of the WIS -------------------------------------------
# 
# df <- eval_forecasts(
#   filtered_data[model %in% regular_models], 
#   metric = "interval_score",
#   summarise_by = c("model", "horizon", "target_type"), 
#   sd = TRUE, 
#   quantile = 0.5
# ) 
# 
# 
# df <- df[, .(model, horizon, target_type, 
#              `WIS - mean` = interval_score, 
#              `WIS - median` = interval_score_0.5, 
#              `WIS - sd` = interval_score_sd)] %>%
#     dcast(... ~ ., value.var = c("WIS - mean", "WIS - median", "WIS - sd"))
# 
# setcolorder(df, c("target_type", "horizon", "model"))
# 
# 
# table <- df[order(target_type, horizon, model)] %>%
#   score_table()
# 
# save_kable(table, 
#            file = here("analysis", "plots", "table-wis-sd-median.png"))

@@

# for horizon <= 2
labels <- df[horizon <= 2][order(target_phase, horizon)]$target_phase
table <- score_table(df[horizon <= 2, !"target_type"], group = FALSE) %>%
  pack_rows(index = table(labels), indent = FALSE)
save_kable(table, 
           file = here("analysis", "plots", "table-phases-2.png"))

# for horizon > 2
labels <- df[horizon > 2][order(target_phase, horizon)]$target_phase
table <- score_table(df[horizon > 2, !"target_type"], group = FALSE) %>%
  pack_rows(index = table(labels), indent = FALSE)
save_kable(table, 
           file = here("analysis", "plots", "table-phases-4.png"))

# 
# horizons = 2
# for (tar in targets) {
#   for (i in horizons) {
#     name <- paste0(tar, "-", i)
#     
#     tmp_df <- df[target_phase == tar & horizon == i, 
#                  .(model, interval_score, sharpness,
#                    target_phase, horizon, 
#                    underprediction, overprediction, 
#                    `WIS - median` = interval_score_0.5, 
#                    `WIS - sd` = interval_score_sd)]
#     
#     tables[[name]] <- score_table(tmp_df, 
#                                        group = FALSE, 
#                                        full_width = FALSE) %>%
#       pack_rows(tar, 1, nrow(tmp_df), indent = FALSE)
#     
#     save_kable(tables[[name]],
#                file = here("analysis", "plots",
#                            paste0("table-", gsub(" ", "", name), ".png")))
#   }
# }

@@

```{r fig.height=10, fig.width=15}
# 
# coverage_tables <- eval_forecasts(
#   filtered_data[grepl("2", target)], 
#   summarise_by = c("model", "target_phase", "classification", "range", "horizon"), 
#   compute_relative_skill = TRUE, 
#   baseline = "Baseline"
# )[range %in% c(50, 90), 
#                    .(model, target_phase, coverage, classification, range)] %>%
#   dcast(formula = ... ~ range, value.var = "coverage")
# 
# scores_tables <- eval_forecasts(
#   filtered_data[grepl("2", target)], 
#   summarise_by = c("model", "target_phase", "classification", "horizon"), 
#   compute_relative_skill = TRUE, 
#   baseline = "Baseline"
# )
# 
# tbl <- scores_tables %>%
#   merge(coverage_tables, all.x = TRUE, by = c("model", "target_phase", "classification")) 
# 
#  tbl <- tbl[, lapply(.SD, FUN = function(x) {
#     if (is.numeric(x)) {
#       return(round(signif(x, 3), 2))
#     } else {
#       return(x)
#     }
#   })]
# 
# tables <- list()
# table_dfs <- list()
# targets <- unique(tbl$target_phase)
# 
# for (tar in targets) {
#   tables[[tar]] <- score_table(tbl[target_phase == tar])
# }
# 
# for (tar in targets) {
#   table_dfs[[tar]] <- tbl[target_phase == tar, 
#                           .("Model" = model, 
#                             # "Target" = target_phase,
#                             "WIS" = interval_score,
#                             # "Rel. skill" = scaled_rel_skill, 
#                             "Sharpn." = sharpness, 
#                             "Underp.." = underprediction, 
#                             "Overp." = overprediction)#, 
#                             # "Bias" = bias,
#                             # "50%-Cov." = `50`, 
#                             # "90%-Cov." = `90`)
#                           ]
# }
# 
# 
# # make plots with tables
# library(gridExtra)
# 
# tbl_plt <- grid.arrange(tableGrob(table_dfs[[1]]), 
#              tableGrob(table_dfs[[2]]), 
#              tableGrob(table_dfs[[3]]), 
#              tableGrob(table_dfs[[4]]), 
#              tableGrob(table_dfs[[5]]), 
#              tableGrob(table_dfs[[6]]), 
#              nrow = 2) 
# 
# tbl_plt <- ggplotify::as.ggplot(tbl_plt)
# 
# tbl_plt
# 
# ggsave("plots/distribution_scores_table.png")
# 
# library(gt)
# 
# tbl[, 
#     .("Model" = model, 
#       "Target" = target_phase,
#       "WIS" = interval_score,
#       # "Rel. skill" = scaled_rel_skill, 
#       "Sharpn." = sharpness, 
#       "Underp.." = underprediction, 
#       "Overp." = overprediction)#, 
#     # "Bias" = bias,
#     # "50%-Cov." = `50`, 
#     # "90%-Cov." = `90`)
# ] %>%
#   gt(rowname_col = "Model", 
#      groupname_col = "Target") 
#

@@

aggregate_performance <- function(data, 
                                  x = "horizon",
                                  xlabel = "Forecast horizon",
                                  output_dir = here("analysis", "plots"), 
                                  plotname,
                                  relative = FALSE,
                                  facet_formula = ~ target_type) {

  scores <- eval_forecasts(data, 
                           summarise_by = c("model", "target_type", x), 
                           quantiles = 0.5, 
                           sd = TRUE)

  plot_helper <- function(tmp = scores, xvar = x, y, rel = relative) {
    if (rel) {
      tmp <- tmp %>%
        arrange(target_type, get(x)) %>%
        group_by(target_type, get(x)) %>%
        mutate_at(y, ~ . - .[model == "Hub-ensemble"])
    }

    tmp %>%
      ggplot(aes_string(x = xvar, y = y, group = "model", color = "model")) +
      geom_point() +
      geom_line() +
      facet_wrap(facet_formula, scales = "free_y") +
      theme_minimal() +
      theme(legend.position = "bottom") +
      scale_color_manual(values = coldeaths)
  }

  # mean performance
  p_mean <- 
    plot_helper(scores, y = "interval_score") + 
    labs(y = "mean WIS", x = NULL)
